External Patch Group Prior Guided Internal Orthogonal Dictionary Learning for Real Image Denoising

### 2016-10-22 CVPR2017_Review_Unified.tex
a

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{upgreek}
\usepackage{multirow}
\usepackage{color}

@@

%%%%%%%%% TITLE
\title{External Patch Group Prior Guided Internal Subspace Learning for Real Image Denoising}

@@

\subsection{Internal v.s. External Image Denoising}
For natural images, the internal patch recurrence across multiple scales has been successfully applied in many image restoration problems \cite{Irani2009sr,Irani2013separating,Irani2014deblur,Irani2016dehazing}. These work demonstrate that internal information is enough for many ill-posed problems including denoising additive white Gaussian noise. The rationale is, since the AWGN noise is independent of the original clean images, it will be reduced if the image is scaled to a smaller size. However, the noise in real images is generated mostly from the camera sensors, which is highly complex and signal dependent \cite{crosschannel2016}. Besides, according to the seminar workd of \cite{healey1994radiometric}, the noise in real images has fixed patterns from several main sources. Therefore, we can hardly seperate the complex noise from the signals without the help of external (correct) information of natural clean images. Only using the internal information may be not enough for real image denoising problem. On the other hand, the methods only using the external information may be not adaptive for real noisy images. Recently, the methods of \cite{mlp,csf,chen2015learning} had been proposed to learn, on both internal and external images, a process directly mapping the noisy patches to denoised ones. These discriminatice learning based methods are only effective on additive white Gaussian noise and ineffective on complex and signal dependent noise in real images. This has been shown in Figure \ref{fig1}. In this paper, our goal is to make use of external information to guide the subspace learning of internal PGs for image denoising task.

@@

\section{External Patch Group Prior Guided Internal Subspace Learning}
In this section, we formulate the framework of external Patch Group prior guided internal subspace learning. We first introduce the patch group prior leaning on clean natural RGB images. Then we formulate the external guided internal subspace learning. Finally, we discuss the differences between external subspaces and the corresponding internal subspace.

@@

\subsection{External Prior Guided Internal Subspace Learning}
Given a rael noisy image, we extract noisy PGs from it and save the mean vectors of each PG for recovering. The mean substracted PG is defined as $\mathbf{\overline{Y}}$. To project this PG into a most adaptive subspace, we select the most suitable Gaussian component to it from the PG-GMM trained in previous section. The selection can be done by checking the posterior probability that $\mathbf{\overline{Y}}$ belongs to the $k$th Gaussian component:
\begin{equation}\label{equ3}
P(k|\mathbf{\overline{Y}})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{l})}.
\end{equation}
Since the noise on real images are mostly small when compared to the signals, the covariance matrix of the $k$th component is still $\mathbf{\Sigma}_{k}$. Finally, the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}})$ is selected as the most suitable subspace for $\mathbf{\overline{Y}}$.

Though each PG has been projected into its most suitable subspace, the pre-learned subspace is still too general to represent the noisy PG extracted from the real noisy image. That is, the noisy PGs projected into one cluster can still constisted a subspace which is of lower dimensions than the subspace pre-learned from the external PGs. This can be demonstrated by compare the distribution of external PGs and internal PGs in the same clusters. We randomly select one cluster, and collect the celan PGs extracted from external dataset (Kodak 24 images) and the niosy PGs from the testing image. Since the original PGs are of $3p^{2}$ dimensions, we apply PCA to project the PGs into 2 dimensions for better visualization. The results is shown in Figure \ref{fig}, from which we can see clearly that the projected PGs are mainly in a smaller region of the external PGs, which proves that the internal PGs are only consisted a subspace in a lower dimension than the PGs collected from external subspace. To better and adaptively charactering the internal PGs from the testing image, we need learn a more specific dictionary for noisy PGs assigned into each cluster. For notation simplicity, we ignore the index of subspace $k$. The internal PGs $\mathbf{Y}$ form a subspace which can be obtained by singular value decomposition (SVD),
\begin{equation}\label{equ4}
\mathbf{Y}=\mathbf{U}\mathbf{S}\mathbf{V}^{T}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

\subsection{What does External Prior help the Internal Subspace Learning?}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information.

\subsection{Discussion on Internal Subspace Structure}
The Internal PGs are in fact lying in the subspaces of external PG Spaces. To defend this argument, we compare the distribution of external PGs extracted from clean natural images and real noisy images. For better illumination, we randomly selected a cluster and project the original clean PGs $\mathbf{X}$ onto a 2-D plane. This could be done via $\mathbf{Xp}=\mathbf{U(:.1:2)}^{T}\mathbf{X}$, where $\mathbf{U}$ is the singular vector matrix of that cluster. The noisy PGs $\mathbf{Y}$ assigned in this cluster is also projected into 2-D via $\mathbf{Yp}=\mathbf{U(:.1:2)}^{T}\mathbf{Y}$. The Figure \ref{} reflects the distribution on the 2-D plane of the projected clean PGs from external natural images and the projected noisy PGs from internal image. We can see that the internal noisy PGs are indeed lying in a subspace of the external PGs. Hence, if we directly use the external prior learned from clean PGs, the learned subspaces would be too generative to be suitable for the testing data.

@@

\begin{table}
\label{alg1}
\begin{tabular}{l}
\hline
\textbf{Alg. 1}: External Guided Internal Subspace Denoising
\\
\hline
\textbf{Input:} Noisy image $\mathbf{y}$, PG-GMM model
\\
1. Initialization: $\hat{\mathbf{x}}^{(0)}=\mathbf{y},\mathbf{y}^{(0)}=\mathbf{y}$;
\\
\textbf{for} $t = 1:IteNum$ \textbf{do}
\\
\quad\textbf{for} each PG $\mathbf{Y}$ \textbf{do}
\\
2.\quad Calculate group mean $\boldsymbol{\upmu}_{y}$ and form PG $\mathbf{\overline{Y}}$;
\\
3.\quad Gaussian component selection via (\ref{equ3});
\\
\quad\textbf{end for}
\\
\quad\textbf{for} each Internal Subspace \textbf{do}
\\
4.\quad Internal Subspace Learning by (\ref{equ4});
\\
5.\quad Recover each patch in all PGs via $\hat{\mathbf{x}}_{m}=\mathbf{D}\hat{\boldsymbol{\upalpha}}+\boldsymbol{\upmu}_{y}$;
\\
\quad\textbf{end for}
\\
6. Aggregate the recovered PGs of all subspaces to form
\\
\quad the recovered image $\hat{\mathbf{x}}^{(t)}$;
\\
\textbf{end for}
\\
\textbf{Output:} The recovered image $\hat{\mathbf{x}}^{(IteNum)}$.\\
\hline
\end{tabular}
\end{table}

>>>>

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{upgreek}
\usepackage{multirow}
\usepackage{color}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

%%%%%%%%% TITLE
\title{External Patch Group Prior Guided Internal Orthogonal Dictionary Learning for Real Image Denoising}

@@

\subsection{Internal v.s. External Dictionary Learning}
For natural images, the internal patch recurrence across multiple scales has been successfully applied in many image restoration problems \cite{Irani2009sr,Irani2013separating,Irani2014deblur,Irani2016dehazing}. These work demonstrate that internal information is enough for many ill-posed problems including denoising additive white Gaussian noise. The rationale is, since the AWGN noise is independent of the original clean images, it will be reduced if the image is scaled to a smaller size. However, the noise in real images is generated mostly from the camera sensors, which is highly complex and signal dependent \cite{crosschannel2016}. Besides, according to the seminar workd of \cite{healey1994radiometric}, the noise in real images has fixed patterns from several main sources. Therefore, we can hardly seperate the complex noise from the signals without the help of external (correct) information of natural clean images. Only using the internal information may be not enough for real image denoising problem. On the other hand, the methods only using the external information may be not adaptive for real noisy images. Recently, the methods of \cite{mlp,csf,chen2015learning} had been proposed to learn, on both internal and external images, a process directly mapping the noisy patches to denoised ones. These discriminatice learning based methods are only effective on additive white Gaussian noise and ineffective on complex and signal dependent noise in real images. This has been shown in Figure \ref{fig1}. In this paper, our goal is to make use of external information to guide the subspace learning of internal PGs for image denoising task.

@@

\section{External Patch Group Prior Guided Internal Orthogonal Dictionary Learning}
In this section, we formulate the framework of external Patch Group prior guided internal subspace learning. We first introduce the patch group prior leaning on clean natural RGB images. Then we formulate the external guided internal subspace learning. Finally, we discuss the differences between external subspaces and the corresponding internal subspace.

@@

\subsection{External Prior Guided Internal Orthogonal Dictionary Learning}
Given a rael noisy image, we extract noisy PGs from it and save the mean vectors of each PG for recovering. The mean substracted PG is defined as $\mathbf{\overline{Y}}$. To project this PG into a most adaptive subspace, we select the most suitable Gaussian component to it from the PG-GMM trained in previous section. The selection can be done by checking the posterior probability that $\mathbf{\overline{Y}}$ belongs to the $k$th Gaussian component:
\begin{equation}\label{equ3}
P(k|\mathbf{\overline{Y}})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{l})}.
\end{equation}
Since the noise on real images are mostly small when compared to the signals, the covariance matrix of the $k$th component is still $\mathbf{\Sigma}_{k}$. Finally, the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}})$ is selected as the most suitable subspace for $\mathbf{\overline{Y}}$.

Though each PG has been projected into its most suitable subspace, the pre-learned subspace is still too general to represent the noisy PG extracted from the real noisy image. That is, the noisy PGs projected into one cluster can still constisted a subspace which is of lower dimensions than the subspace pre-learned from the external PGs. This can be demonstrated by compare the distribution of external PGs and internal PGs in the same clusters. We randomly select one cluster, and collect the celan PGs extracted from external dataset (Kodak 24 images) and the niosy PGs from the testing image. Since the original PGs are of $3p^{2}$ dimensions, we apply PCA to project the PGs into 2 dimensions for better visualization. The results is shown in Figure \ref{fig}, from which we can see clearly that the projected PGs are mainly in a smaller region of the external PGs, which proves that the internal PGs are only consisted a subspace in a lower dimension than the PGs collected from external subspace. To better and adaptively charactering the internal PGs from the testing image, we need learn a more specific dictionary for noisy PGs assigned into each cluster. For notation simplicity, we ignore the index of subspace $k$. The internal PGs $\mathbf{Y}$ form a subspace which can be obtained by singular value decomposition (SVD),
\begin{equation}\label{equ4}
\begin{split}
&\min_{\mathbf{D}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e};\mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

\subsection{Optimization with Closed-form Solution}
Similar to the K-SVD \cite{ksvd}, we employ an alternating iterative framework to solve the optimization problem \ref{equ4}. In fact, we initialize the orthogonal dictionary as $\mathbf{D}^{(0)}$ and for $t=0, 1, ..., T-1$, alternatively do

\textbf{Updating Weighted Sparse Coding}: given the initialization orthogonal dictioanry $\textbf{D}_{i}^{(t)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}
\textbf{A}^{(t)}:=\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e};\mathbf{D}_{i}^{(t)}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\end{equation}

\textbf{Updating Orthogonal Dictionary}: given the sparse coefficients $\textbf{A}^{(0)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}
\begin{split}
\textbf{D}_{i}^{(t+1)}:=
&
\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e};\mathbf{D}_{i}]\mathbf{A}^{(t)}\|_{F}^{2}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}

\subsection{Discussion on External Prior and Internal Orthogonal Dictionary Learning}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information.

The Internal PGs are in fact lying in the subspaces of external PG Spaces. To defend this argument, we compare the distribution of external PGs extracted from clean natural images and real noisy images. For better illumination, we randomly selected a cluster and project the original clean PGs $\mathbf{X}$ onto a 2-D plane. This could be done via $\mathbf{Xp}=\mathbf{U(:.1:2)}^{T}\mathbf{X}$, where $\mathbf{U}$ is the singular vector matrix of that cluster. The noisy PGs $\mathbf{Y}$ assigned in this cluster is also projected into 2-D via $\mathbf{Yp}=\mathbf{U(:.1:2)}^{T}\mathbf{Y}$. The Figure \ref{} reflects the distribution on the 2-D plane of the projected clean PGs from external natural images and the projected noisy PGs from internal image. We can see that the internal noisy PGs are indeed lying in a subspace of the external PGs. Hence, if we directly use the external prior learned from clean PGs, the learned subspaces would be too generative to be suitable for the testing data.

@@

\begin{table}
\label{alg1}
\begin{tabular}{l}
\hline
\textbf{Alg. 1}: External Guided Internal Dictionary Learning
\\
\hline
\textbf{Input:} Noisy image $\mathbf{y}$, PG-GMM model
\\
1. Initialization: $\hat{\mathbf{x}}^{(0)}=\mathbf{y},\mathbf{y}^{(0)}=\mathbf{y}$;
\\
\textbf{for} $t = 1:IteNum$ \textbf{do}
\\
\quad\textbf{for} each PG $\mathbf{Y}$ \textbf{do}
\\
2.\quad Calculate group mean $\boldsymbol{\upmu}_{y}$ and form PG $\mathbf{\overline{Y}}$;
\\
3.\quad Gaussian component selection via (\ref{equ3});
\\
\quad\textbf{end for}
\\
\quad\textbf{for} each Internal Subspace \textbf{do}
\\
4.\quad Internal Subspace Learning by (\ref{equ4});
\\
5.\quad Recover each patch in all PGs via $\hat{\mathbf{x}}_{m}=\mathbf{D}\hat{\boldsymbol{\upalpha}}+\boldsymbol{\upmu}_{y}$;
\\
\quad\textbf{end for}
\\
6. Aggregate the recovered PGs of all subspaces to form
\\
\quad the recovered image $\hat{\mathbf{x}}^{(t)}$;
\\
\textbf{end for}
\\
\textbf{Output:} The recovered image $\hat{\mathbf{x}}^{(IteNum)}$.\\
\hline
\end{tabular}
\end{table}
